---
title: 'ILSE: installation and simulated example'
author: "Wei Liu"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ILSE: installation and simulated example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```



## Install the ILSE
This vignette provides an introduction to the R package `ILSE`, where the function `ILSE` implements the model `ILSE`, Linear Regression by Iterative Least Square Estimation (ILSE) When Covariates Include Missing Values.
The package can be installed with the command from [Github](https://github.com/feiyoung/ILSE):

`library(remotes)`

`remotes::install_github("feiyoung/ILSE")`

or install from [CRAN](https://cran.r-project.org/)

`install.packages("ILSE")`

The package can be loaded with the command:
```{r}
library("ILSE")
```

## Fit linear model using simulated data
First, we generate the data with  homogeneous  normal variables.
```{r  eval = FALSE}
  n <- 100
  p <- 6
  X <- MASS::mvrnorm(n, rep(0, p), cor.mat(p, rho=0.5))
  beta0 <- rep(c(1,-1), times=3)
  Y <- -2+ X %*% beta0 + rnorm(n, sd=1)
```

### A special case: without missing values

Then, we fit the linear regression model without missing values based on ILSE.
```{r eval = FALSE}
ilse1 <- ilse(Y~X)
print(ilse1)
```

We can also create a (data.frame) object as input for ILSE.

```{r eval = FALSE}
dat <- data.frame(Y=Y, X=X)
ilse1 <- ilse(Y~., data=dat)
print(ilse1)
Coef(ilse1) # access the coefficients
Fitted.values(ilse1)
Residuals(ilse1)

```

Check the significant variables by bootstratp.
```{r eval = FALSE}
s1 <- summary(ilse1)
s1
```

### Handle data with missing values
First, we randomly remove some entries in X.
```{r eval = FALSE}
mis_rate <- 0.3
set.seed(1)
na_id <- sample(1:(n*p), n*p*mis_rate)
Xmis <- X
Xmis[na_id] <- NA
ncomp <- sum(complete.cases(Xmis))
message("Number of complete cases is ", ncomp, '\n')
```

Second, we use *lm* to fit linear regression model based on complete cases, i.e., CC analysis. We can not detect any siginificant covariates.

```{r eval = FALSE}
lm1 <- lm(Y~Xmis)
s_cc <- summary.lm(lm1)
s_cc
```

Third, we use *ILSE* to fit the linear regression model based on all data.
```{r eval = FALSE}
ilse2 <- ilse(Y~Xmis, data=NULL, infor_output=T)
print(ilse2)
```

Fourth, *Bootstrap* is applied to evaluate the standard error and p-values of each coefficients estimated by ILSE. We observe four significant coefficients. 
```{r eval = FALSE}
s2 <- summary(ilse2, Nbt=20)
s2
```


In *ILSE* package, we also provide Full Information Maximum Likelihood for Linear Regression *fimlreg*. We show how to use it to handle the above missing data.
```{r eval = FALSE}
fimllm <- fimlreg(Y~Xmis)
print(fimllm)

```

We also use *bootstrap* to evaluate the standard error and p-values of each coefficients estimated by ILSE. We observe only one significant coefficients. 
```{r eval = FALSE}
s_fiml <- summary(fimllm, Nbt=20)
s_fiml
```

### Visualization 
We visualize the p-vaules of each methods , where red line denotes 0.05 in y-axis and blue line 0.1 in y-axis.
```{r eval = FALSE}
pMat <- cbind(CC=s_cc$coefficients[,4], ILSE=s2[,4], FIML=s_fiml[,4])
library(ggplot2)
df1 <- data.frame(Pval= as.vector(pMat[-1,]),
                    Method =factor(rep(c('CC', "ILSE", "FIML"),each=p)),
                    covariate= factor(rep(paste0("X", 1:p), times=3)))
ggplot(data=df1, aes(x=covariate, y=Pval, fill=Method)) + geom_bar(position = "dodge", stat="identity",width = 0.5) + geom_hline(yintercept = 0.05, color='red') + geom_hline(yintercept = 0.1, color='blue')
```


## Session information
```{r}
sessionInfo()
```
